---
title: "Random Samples"
author: "Patrick Robotham"
date: "January 23, 2018"
output: html_document
---

# Simulations

## The Classical Statistical Set Up.

For a random variable X, obtain n random samples of X.

Let's call X[n] a "sample vector" of X.
(e.g. `rnorm(100)` is a sample vector of N(0,1) (of length 100)

"Random samples" are also known as "deviates"

https://www.amazon.com/Million-Random-Digits-Normal-Deviates/dp/0833030477

Point: Sample vectors are data! We can manipulate them using Pandas, Dplyr, or Excel.

http://probabilitymanagement.org/sip-math.html

## Intervals

One thing we often want to do with our sample vectors is summarise them as an interval.

Many different options: `?point_interval` , `binom.confint`

http://staff.math.su.se/hoehle/blog/2017/06/18/quantiles.html

## Averages

Often we describe sample vectors with reference to their 'average' (usually mean).
This is sometimes justified by the law of large numbers.
Population variability is disregarded.

"More recently, a failure to appreciate the flaw led to $2 billion in property damage in North Dakota. In 1997, the U.S. Weather Service forecast that North Dakota’s rising Red River would crest at 49 feet. Officials in Grand Forks made flood management plans based on this single figure, which represented an average. In fact, the river crested above 50 feet, breaching the dikes, and unleashing a flood that forced 50,000 people from their homes."

https://hbr.org/2002/11/the-flaw-of-averages


## Independence
Example: mean(1:9)^2  != mean((1:9)^2))
Jutstification: Locality.


GFC was arguably caused by a false presumption of independence (low correlation)
https://www.businessinsider.com.au/bubble-derivatives-otc-2010-5

## The Statistician/Manager Axis

   Raw Data - Distribution - Interval - Average - Category
<- Statistician   -----------------------          Executive ->

# Stan/Code
## Generated Quantities
Stan handles simulation problem

At present have to fit model before generating quantities (slow).

https://github.com/stan-dev/stan/issues/2432

## Extract

Extract returns a list of arrays.

## Dataframes

Your model's samples should be represented as a table!

Advantages:
* Can use pandas / dplyr

* Many useful plotting libraries expect a table (e.g. ggplot)

* Can export to csv (and view in Excel)

* Can export to Database (to power code)


## Tidybayes
An R package `tidybayes` exists to solve this
https://github.com/mjskay/tidybayes


Example taken from  
https://github.com/stan-dev/stancon_talks/tree/master/2017/Contributed-Talks/01_auerbach


Old and broken
```{r load fitted model}
load(file = "fit.rds")
```
```{r old_postprocessing}
wthn <- c("COND_e","CITY_e","YEAR_e","SLIM_e",
          "SIGN_e","LGHT_e","BLTE_e","TFFC_e")
coefs <- extract(fit,pars=wthn)
coef_ggplot <- data.frame(wthn = character(),
                          numb = character(),
                          coef_mean=numeric(),
                          upper50 = numeric(),
                          lower50 = numeric(),
                          upper95 = numeric(),
                          lower95 = numeric())
for(var in seq_along(coefs)){
  coefs_temp <- coefs[[var]]
  coef_ggplot_temp <- data.frame(wthn = wthn[var], 
                                 numb = paste0(wthn[var],1:ncol(coefs_temp)),
                                 coef_mean=apply(coefs_temp,2,mean))
  coef_ggplot_temp$upper50 <- apply(coefs_temp,2,quantile,probs=.75)
  coef_ggplot_temp$lower50 <- apply(coefs_temp,2,quantile,probs=.25)
  coef_ggplot_temp$upper95 <- apply(coefs_temp,2,quantile,probs=.975)
  coef_ggplot_temp$lower95 <- apply(coefs_temp,2,quantile,probs=.025)
  coef_ggplot <- rbind(coef_ggplot,coef_ggplot_temp)
}
coef_ggplot
```

New Hotness

```{r new_postprocessing}
library(tidybayes)
library(tidyverse)
coef_ggplot_new <- fit %>% 
  gather_samples(`.*_e`[index],regex = TRUE) %>% 
  group_by(term, index) %>%
  summarise(
    coef_mean = mean(estimate),
    upper50 = quantile(estimate,0.75),
    lower50 = quantile(estimate,0.25),
    upper95 = quantile(estimate,0.975),
    lower95 = quantile(estimate,0.025)
  ) 
coef_ggplot_new
```



## Common tasks
* Compute a marginal distribution  X[I,J] | J=j
Use `m %>% gather_samples %>% filter(index2 == j)`
* Compute an aggregate distribution sum_I X[I]
Use `m %>% gather_samples %>% group_by(term, sample) %>% summarise(total = sum(estimate))`

* Get confidence intervals / averages
Use `m %>% spread_samples %>% summarise_all(mean)`

